### 2020-04-15 Thesis-Chapter1-Introduction.tex
Write some more; add international EA resource

\Acl{EA} is a wicked problem. In encryption policy, \ac{EA} is alternative means of decryption intended for law
enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
resistance to a clear solution, the debate on \ac{EA} refuses to end. This paper will not attempt to end the debate, but
to inform it with additional structure.

@@

The conflict at the heart of the encryption and \ac{EA} is that the same cryptographic and design principles that
underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
communications and storage provides clear benefits and should be welcomed by all. In a world where information security
is frightfully poor yet increasingly important, strong encryption's necessity can not be understated. However, the
privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing. In a
society that cares about bringing wrongdoers to justice, this risk should not be ignored.

The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
access encrypted data to perform their duties. Human rights groups and technical leaders counter that a weakened
encryption environment would fatally compromise privacy and security as we know it. As the debate continues, sides have
become entrenched, and both sides have engaged in disingenuous attacks against their perceived opponents.

@@

This paper serves to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
brief U.S.-focused history of encryption regulation, technology, and conflict, it introduces argument maps and threat
models using data flow diagrams. In the following chapters it analyzes exceptional access arguments in detail and
creates a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and
\acl{DIM}. The paper concludes with a set of policies and principles to apply to the problem in light of valid arguments
and threats.

>>>>

\Acl{EA} is a wicked problem. In encryption policy, \ac{EA} is alternative means of decryption intended for law
enforcement use. Characterized by a dynamic technological environment, disagreement about underlying values, and
resistance to a clear solution, the debate on \ac{EA} refuses to end. This thesis will not attempt to end the debate,
but to inform it with additional structure.

@@

The conflict at the heart of the encryption and \ac{EA} debate is that the same cryptographic and design principles that
underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
communications and storage provides clear benefits and should be welcomed by all. In a world where information security
is frightfully poor yet increasingly important, strong encryption's necessity can not be understated. However, the
privacy afforded by certain encryption technologies hampers law enforcement investigations and hides wrongdoing. 
% TODO: citation
In a
society that cares about bringing wrongdoers to justice, this risk should not be ignored.

The conversation over encryption is in a stalemate. Governments and law enforcement agencies frequently cite the need to
access encrypted data to perform their duties. 
% TODO: citation
Human rights groups and technical leaders counter that a weakened
encryption environment would fatally compromise privacy and security as we know it. 
% TODO: citation
As the debate continues, sides have
become entrenched, and both sides have engaged in disingenuous attacks against their perceived opponents.

@@

This thesis aims to clarify the arguments composing debate and the nature of the technical problem. After reproducing a
brief U.S.-focused history of encryption regulation, technology, and conflict, I introduce argument maps and threat
models using data flow diagrams. In the following chapters, I analyze exceptional access arguments in detail and create
a multi-level threat model, broken down by the significantly different challenges posed by \acl{DAR} and \acl{DIM}. The
thesis concludes with a set of policies and principles to apply to the problem in light of valid arguments and threats.

% TODO: use first person voice? Originally had wording like "this thesis introduces", but Kalafut says not to be afraid
% of "I". Could also use passive voice.

### 2020-04-15 Thesis-Chapter2-Background.tex
Write some more; add international EA resource

There are two major cryptographic protocol families--symmetric and asymmetric. In traditional \ac{sym-crypto},
encryption and decryption are performed with the same key, and both parties must have this key in order to communicate
securely. In \ac{asym-crypto}, encryption and decryption are performed with separate, paired keys, called the public key
and the private key. The innovation of \ac{public-key-crypto} is that the public key is not secret, and using this
technique, two parties can communicate securely without requiring an a previously agreed upon shared secret.

@@

The following sections list laws which directly or indirectly affect use encryption. \Ac{EA} is an international
problem, as digital technology has little respect for borders. This paper is U.S.-focused, but briefly touches on the
regulatory environment of other countries.

@@

\lawsstart
    \law{1789}{\acf{AWA}}{congress_1789}{

@@

@@

    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{}

% \cite{wolff_computer_2016}

    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}{}

    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}{}

    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{}

    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{}

    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{}

    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{}

@@

\lawsstart
    \law{1993}{The \ac{clipperchip}}{press_1993}{}
% \cite{thompson_2015}
    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}{}
    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{}
% mix of SE protection and EA guidelines?
    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}{ashcroft_s2067_1998}{}
% like above?
    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{}
    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{}
% ^^ these two introduced by same rep on same day
    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}{}
    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}{}
% federal SE protections
    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}{}
% state SE protections
    \law{2020}{The EARN IT Act}{graham_s3398_2020}{}
\lawsend

@@

In order to advance the \ac{EA} debate, it helps to organize all the arguments used in a cohesive manner. One way to
organize arguments is to create argument maps. Argument maps are graphical representations of the logical structures and
relationships between statements, premises, and conclusions. Argument maps are an outgrowth of the same research by
Horst Rittel that originated the idea of \acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break
down \acp{wicked-prob} and document the reasoned approach that goes into solving them \cite{kunz_issues_1970}. Interest
in \ac{IBIS} has waxed and waned. Some software tools have been developed, including the now-defunct gIBIS
\cite{conklin_gibis_1988} and Compendium \cite{dutoit_hypermedia_2006} tools designed to produce graphical
representations. Argdown is a recent tool that generates argument maps from structures specified in a Markdown-based
language \cite{voigt_argdown_2018}.

This paper uses Argdown to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo} shows the basic
structure of arguments. Statements are positions on issues. Some statements are based on values that cannot be addressed
through argument. Statements and arguments may support or attack one another, and arguments support or undercut other
arguments.

@@

\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
external entities. \Acp{DFD} well suited for threat modeling because security vulnerabilities tend to follow data flow,
not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is the
primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. This paper creates \acp{DFD} in this style
to analyze the threats already present and those that \ac{EA} would introduce.

>>>>

There are two major cryptographic protocol families--symmetric and asymmetric. In traditional \ac{sym-crypto},
encryption and decryption are performed with the same key, and both parties must have this key in order to communicate
securely. In \ac{asym-crypto}, also called \ac{public-key-crypto}, encryption and decryption are performed with two
paired keys, called the public key and the private key. The innovation of \ac{public-key-crypto} is that the public key
is not secret, and using this technique, two parties can communicate securely without requiring an a previously agreed
upon shared secret.

@@

The following sections list laws which directly or indirectly affect use encryption. \Ac{EA} is an international
problem, as digital technology has little respect for borders. This thesis is U.S.-focused, but briefly touches on the
regulatory environment of other countries.

@@

\lawsstart

    \law{1789}{\acf{AWA}}{congress_1789}{

@@

% IDEA: go into
%  Wiretap Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_119.html)
%  Stored Communications Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_I_20_121.html)
%  Pen Register Act (http://www.law.cornell.edu/uscode/usc_sup_01_18_10_II_20_206.html)
% etc?

@@

    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{

% FIXME: is this actually related enough to keep?

Criminalized ``knowingly access[ing] a computer without authorization or exceeding authorized access'' and thereby
violating the confidentiality, integrity, or availability of the system. While presumably intended to target hacking,
the \acs{CFAA} has been used to prosecute a host of crimes, and suffers from dated language;
% --what does it mean to ``access'' a ``computer'' loading a simple webpage may involve dozens of devices--
whether violating the terms of service constitutes ``exceeding authorized access'' remains an open question
\cite{wolff_computer_2016}. Importantly for \ac{EA}, the law includes carve-outs for law enforcement and intelligence
activities.

    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}{

% Amends Wiretap Act, Stored Communications Act, Pen Register Act. Deep history, but picking up here.

    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}{

    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{

    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{

    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{

    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{

@@

\lawsstart

    \law{1993}{The \ac{clipperchip}}{press_1993}{

}
% \cite{thompson_2015}

    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}{

    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{

% mix of SE protection and EA guidelines?

    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}
        {ashcroft_s2067_1998}{

% like above?

    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{

    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{

% ^^ these two introduced by same rep on same day

    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}{

    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}{

% federal SE protections

    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}{

% state SE protections

    \law{2020}{The EARN IT Act}{graham_s3398_2020}{

@@

In order to advance the \ac{EA} debate, it helps to organize all the arguments used in a cohesive manner. One way to
organize arguments is to create argument maps. Argument maps are graphical representations of the logical structures and
relationships between statements, premises, and conclusions. Argument maps are an outgrowth of the same research by
Horst Rittel that originated the idea of \acp{wicked-prob}. In 1970 his research group developed \acp{IBIS} to break
down \acp{wicked-prob} and document the reasoned approaches that go into solving them \cite{kunz_issues_1970}. Interest
in \ac{IBIS} has waxed and waned. Some software tools have been developed, including the now-defunct gIBIS
\cite{conklin_gibis_1988} and Compendium \cite{dutoit_hypermedia_2006} tools designed to produce graphical
representations. Argdown is a recent tool that generates argument maps from structures specified in a Markdown-based
language \cite{voigt_argdown_2018}.

% TODO: reference forward chapter
Argdown is used in this thesis to analyze the arguments used in the \ac{EA} debate. \myfig{fig-args-demo} shows the
basic structure of arguments. Statements are positions on issues. Some statements are based on values that cannot be
addressed through argument. Statements and arguments may support or attack one another, and arguments support or
undercut other arguments.

@@

\Acp{DFD} are one tool designed to address these questions. \Acp{DFD} feature processes, data flows, data stores, and
external entities. \Acp{DFD} well suited for threat modeling because security vulnerabilities tend to follow data flow,
not control flow \cite{shostack_threat_2014}. They are particularly well suited for \ac{EA}, as data privacy is the
primary concern. \myfig{fig-dfd-demo} shows the basic elements of a \ac{DFD}. In this thesis I create \acp{DFD} in this
style to analyze the threats already present and those that \ac{EA} would introduce.
% TODO: reference forward chapter
